Practicle Machine Learning Project
=================================

## Background:  

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. In this test, people were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

The data is available at: http://groupware.les.inf.puc-rio.br/har (section on the Weight Lifting Exercise Dataset). 
The training set data is available at: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The testing set is available at: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

## Data loading:

```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(fileUrl, destfile = "training.csv", method = "curl")
fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(fileUrl, destfile = "test.csv", method = "curl")
list.files()
train_data <- read.table("training.csv", sep = ",", header = TRUE)
test_data <- read.table("test.csv", sep = ",", header = TRUE)

names(train_data)
head(test_data)

dim(train_data)
dim(test_data)
```

## Cleaning data (Removing NA's):

```{r}
clean_train <- train_data[, colSums(is.na(train_data)) == 0]
clean_test <- test_data[, colSums(is.na(test_data)) == 0]
```

## Discarding non useful variables:

```{r}
n_c_train_1 <- clean_train[, !grepl("X|user_name|timestamp|window", colnames(clean_train))]
n_c_test_1 <- clean_test[, !grepl("X|user_name|timestamp|window", colnames(clean_test))]

n_c_train_2 <- n_c_train_1[, !grepl("^max|^min|^ampl|^var|^avg|^stdd|^ske|^kurt", colnames(n_c_train_1))]
n_c_test_2 <- n_c_test_1[, !grepl("^max|^min|^ampl|^var|^avg|^stdd|^ske|^kurt", colnames(n_c_test_1))]
dim(n_c_train_2)
dim(n_c_test_2)
```

## Cross validation (70% training and 30% validation):

```{r}
set.seed(23222)
library(caret)
in_train <- createDataPartition(y = n_c_train_2$classe, p = 0.7, list = FALSE)
train <- n_c_train_2[in_train, ]
train_valid <- n_c_train_2[-in_train, ]
n_c_test <- n_c_test_2
```
## Ploting data:

```{r}
train_corr <- cor(train[, -53])
heatmap(train_corr)
library(corrplot)
corrplot(train_corr, method = "color")
```

Highly correlated variables
```{r}
h_c <- abs(train_corr)
diag(h_c) <- 0
h_corr <- which(h_c > 0.8, arr.ind = TRUE)
for (i in 1:nrow(h_corr)) {
  print(names(train)[h_corr[i, ]])
}
```

## PCA for combination of predictors to have reduced number of predictors and reduced noise:

```{r}
train_pca <- preProcess(train[, -53], method  = "pca", thresh = 0.95) # 0.9
train_pca_1 <- predict(train_pca, train[, -53])
valid_pca <- predict(train_pca, train_valid[, -53])
test_pca <- predict(train_pca, n_c_test[, -53])
print(train_pca)
```

## Random forests for cross validation method:

```{r}
mod <- train(train$classe ~ ., method = "rf", data = train_pca_1, trControl = trainControl(method = "cv", 5)) # 5
mod
varImpPlot(mod$finalModel, sort = TRUE)
```

## Cross validation:

```{r}
pred_valid <- predict(mod, valid_pca)
confusionMatrix(train_valid$classe, pred_valid)
OoSE <- 1 - as.numeric(confusionMatrix(train_valid$classe, pred_valid)$overall[1])
pred_test <- predict(mod, test_pca)
```

